---
title: "Importing and Using FRM Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Importing and Using FRM Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(AirVizR)
library(readxl)
library(lubridate)
library(magrittr)
```

# Importing local Federal Reference Monitor (FRM) data
The following example demonstrates how to wrangle local FRM data.
It is based on the data format of the [Oregon Department of Environmental Quality (DEQ)](https://oraqi.deq.state.or.us/home/map) "Portland SE Lafayette" station, which you can download from their ["single station report" tool](https://oraqi.deq.state.or.us/report/SingleStationReport).
Other FRM data can be imported in a similar manner, but might require customization.

---

## Inputs

```{r inputs}
# File path to the file of interest
input_frm_path <- "data-raw/frm_july.xlsx"

## Optional grouping settings (which will NOT impact raw data output, but can be useful when graphing!)

## DATE GROUPING
# Run date grouping? If set to "FALSE", the date inputs below will not matter (but will still appear in the environment)
run_date_grouping <- TRUE
# If TRUE, specify the groupings below as respective items in their lists
# (i.e. the first item in each list is one "group", the second item in each list is another "group", etc.)
# Date grouping categories:
input_date_tags <- c("Before", "Independence Day", "After")
# Start dates (in a list, "YYYY-MM-DD")
input_date_starts <- c("2020-07-01", "2020-07-04", "2020-07-05")
# End dates (in a list, "YYYY-MM-DD")
input_date_ends <- c("2020-07-03", "2020-07-04", "2020-07-07")

## HOUR GROUPING
# Run hour grouping? If set to "FALSE", the hour inputs below will not matter (but will still appear in the environment)
run_hour_grouping <- TRUE
# If TRUE, specify the groupings below as respective items in their lists
# (i.e. the first item in each list is one "group", the second item in each list is another "group", etc.)
# Hour grouping categories:
input_hour_tags <- c("Morning", "Afternoon", "Evening", "Night")
# Start hours (in a list, using full hours in 24 hour format)
input_hour_starts <- c(5, 12, 17, 21)
# End hours will be automatically generated; no need to add "end" hours
```

```{r load_raw_data}
# The following is location data for the DEQ FRM in Portland
frm_raw_meta <- data.frame(
  label = "DEQ SE Lafayette",
  location = factor("FRM"), # The device is outside but this will label it separately
  latitude = 45.496641,
  longitude = -122.602877,
  site_id = "41-051-0080",
  timezone = "America/Los_Angeles", # Use OlsonNames() to see valid inputs.
  flag_highValue = FALSE # Being an FRM, it is not expected to be reporting poorly
)

raw_frm <- read_frm()
```


```{r wrangle_data}
frm_meta <- wrangle_meta(frm_raw_meta)
frm_full <- wrangle_frm(raw_frm, frm_meta)

frm_hourly <- group_stad(frm_full, by_day = TRUE, by_hour = TRUE) %>% apply_hour_tags()
frm_daily <- group_stad(frm_full, by_day = TRUE, by_hour = FALSE) %>% apply_date_tags()
frm_diurnal <- group_stad(frm_full, by_day = FALSE, by_hour = TRUE) %>% apply_hour_tags()

if (run_date_grouping == TRUE){
  frm_full <- apply_date_tags(frm_full)
}
if (run_hour_grouping == TRUE) {
  frm_full <- apply_hour_tags(frm_full)
}
```

```{r start_end_dates}
if (exists("input_startdate") == FALSE) {
  input_startdate <- min((frm_full %>% column_dt("date"))$date)
}
if (exists("input_enddate") == FALSE) {
  input_enddate <- max((frm_full %>% column_dt("date"))$date)
}
```

```{r eclean_applying}
# OPTIONAL
# Cleaning the global environment
remove(run_date_grouping, run_hour_grouping)
remove(list = c(ls(pattern = "input_(hour|date|frm)_")))
# It is recommended to keep raw_meta and raw_data since it takes the longest to load,
# as well as the input start & end dates (since the diurnal set does not contain date information)
```

---

## Next steps

Now we're ready to visualize! See the [`visualize-data` vignette](visualize-data.html) for instructions.
