---
title: "Importing and Using FRM Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Importing and Using FRM Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(AirVizR)
library(readxl)
library(lubridate)
```

# Importing local Federal Reference Monitor (FRM) data
The following example demonstrates how to wrangle local FRM data.
It is based on the data format of the [Oregon Department of Environmental Quality (DEQ)](https://oraqi.deq.state.or.us/home/map) "Portland SE Lafayette" station, which you can download from their ["single station report" tool](https://oraqi.deq.state.or.us/report/SingleStationReport).
Other FRM data can be imported in a similar manner, but might require customization.

---

## Inputs

```{r inputs}
# File path to the file of interest
input_frm_path <- "data-raw/frm_july.xlsx"

## Optional grouping settings (which will NOT impact raw data output, but can be useful when graphing!)

## DATE GROUPING
# Run date grouping? If set to "FALSE", the date inputs below will not matter (but will still appear in the environment)
run_date_grouping <- TRUE
# If TRUE, specify the groupings below as respective items in their lists
# (i.e. the first item in each list is one "group", the second item in each list is another "group", etc.)
# Date grouping categories:
input_date_tags <- c("Before", "Independence Day", "After")
# Start dates (in a list, "YYYY-MM-DD")
input_date_starts <- c("2020-07-01", "2020-07-04", "2020-07-05")
# End dates (in a list, "YYYY-MM-DD")
input_date_ends <- c("2020-07-03", "2020-07-04", "2020-07-07")

## HOUR GROUPING
# Run hour grouping? If set to "FALSE", the hour inputs below will not matter (but will still appear in the environment)
run_hour_grouping <- TRUE
# If TRUE, specify the groupings below as respective items in their lists
# (i.e. the first item in each list is one "group", the second item in each list is another "group", etc.)
# Hour grouping categories:
input_hour_tags <- c("Morning", "Afternoon", "Evening", "Night")
# Start hours (in a list, using full hours in 24 hour format)
input_hour_starts <- c(5, 12, 17, 21)
# End hours will be automatically generated; no need to add "end" hours
```

```{r location_data}
# The following is location data for the DEQ FRM in Portland
frm_meta <- data.frame(
  label = "DEQ SE Lafayette",
  location = factor("FRM"), # The device is outside but this will label it separately
  latitude = 45.496641,
  longitude = -122.602877,
  site_id = "41-051-0080",
  timezone = "America/Los_Angeles", # Use OlsonNames() to see valid inputs.
  flag_highValue = FALSE # Being an FRM, it is not expected to be reporting poorly
)
```

```{r load_data}
raw_frm <- read_frm()
```

```{r}
frm_full <- wrangle_frm(raw_frm, frm_meta)

frm_hourly <- group_stad(frm_full, by_day = TRUE, by_hour = TRUE)
frm_daily <- group_stad(frm_full, by_day = TRUE, by_hour = FALSE)
frm_diurnal <- group_stad(frm_full, by_day = FALSE, by_hour = TRUE)

if (run_date_grouping == TRUE){
  frm_full <- apply_date_tags(frm_full)
}
if (run_hour_grouping == TRUE) {
  frm_full <- apply_hour_tags(frm_full)
}
```

```{r start_end_dates}
if (exists("input_startdate") == FALSE) {
  input_startdate <- min((frm_full %>% column_dt("date"))$date)
}
if (exists("input_enddate") == FALSE) {
  input_enddate <- max((frm_full %>% column_dt("date"))$date)
}
```

```{r eclean_applying}
# OPTIONAL
# Cleaning the global environment
remove(run_date_grouping, run_hour_grouping)
remove(list = c(ls(pattern = "input_(hour|date|frm)_")))
# It is recommended to keep raw_meta and raw_data since it takes the longest to load,
# as well as the input start & end dates (since the diurnal set does not contain date information)
```


```{r data_import}

# Creating a list of variable names
deq_names <- read_excel(input_deq_path,
                        sheet = 1, # Selecting sheet 1
                        skip = 2, # Skipping the empty rows at the top
                        n_max = 0) %>% names() # Selecting 0 observations, and extracting the names only

# All DEQ data aside from time stamps is numeric
# To avoid R from identifying columns with lots of missing data as logical, all columns except date time are set to numeric:
deq_coltypes <- ifelse(grepl("Date Time", deq_names), "text", "numeric")

# Creating a data frame of the data
raw_deq <- read_excel(input_deq_path,
                      sheet = 1, # Selecting sheet 1
                      skip = 4, # Skipping the topmost rows naming the variables, etc
                      na = "----", # NAs are reported using this symbol in the Excel file
                      col_names = deq_names, # Setting column names to be those listed above
                      col_types = deq_coltypes) # Setting column types

# Data frame of units, for reference if needed.
deq_units <- read_excel(input_deq_path,
                        sheet = 1, # Selecting sheet 1
                        skip = 2, # Skipping the empty rows at the top
                        n_max = 1) # Selecting 1 "observation" (next row)
```

```{r eclean_read}
## OPTIONAL
remove(deq_names, deq_coltypes, deq_units, input_deq_path)
```


```{r data_wrangling}
data_deq <- raw_deq %>% 
  mutate(
    # Oregon DEQ reports time stamps in Pacific Standard Time, meaning DST is NOT accounted for.
    # Using Etc/GMT+8 here will mark the time stamps as the appropriate -8 ("PST")
    datetime_standard = parse_date_time(`Date Time`, "%H:%M %m/%d/%Y", tz = "Etc/GMT+8"),
    # The following will adjust time stamps for DST appropriately:
    `Date Time` = with_tz(datetime_standard, tzone = "America/Los_Angeles")
  ) %>% 
  rename(datetime = `Date Time`) %>% 
  # Renaming variables based on inputs at the top
  plyr::rename(warn_missing = FALSE, input_deq_renames) %>% 
  mutate(site_id = data_deq_meta$site_id) %>% 
  # Selecting only renamed variables in addition to time stamps
  select(c(site_id, datetime, intersect(all_of(input_deq_renames), names(.)), datetime_standard))

# Converting deg Ç (units in which DEQ reports) to deg F (units in which PA reports)
if ("temperature_c" %in% colnames(data_deq) == TRUE & "temperature" %in% colnames(data_deq) == FALSE) {
  data_deq <- data_deq %>% 
    mutate(temperature = unit_convert(celsius = temperature_c))
}
```

```{r data_applying}
data_deq_hourly <- group_stad(data_deq, by_day = TRUE, by_hour = TRUE)
data_deq_daily <- group_stad(data_deq, by_day = TRUE, by_hour = FALSE)
data_deq_diurnal <- group_stad(data_deq, by_day = FALSE, by_hour = TRUE)

if (run_date_grouping == TRUE){
  data_deq <- data_deq %>% apply_date_tags()
}
if (run_hour_grouping == TRUE) {
  data_deq <- data_deq %>% apply_hour_tags()
}
```

```{r start_end_dates}
if (exists("input_startdate") == FALSE) {
  input_startdate <- as.character(min((data_deq %>% column_dt("date"))$date))
}
if (exists("input_enddate") == FALSE) {
  input_enddate <- as.character(max((data_deq %>% column_dt("date"))$date))
}
```

```{r eclean_applying}
# OPTIONAL
# Cleaning the global environment
remove(run_date_grouping, run_hour_grouping)
remove(list = c(ls(pattern = "input_(hour|date)_")))
# It is recommended to keep raw_meta and raw_data since it takes the longest to load,
# as well as the input start & end dates (since the diurnal set does not contain date information)
```

---

## Next steps

Now we're ready to visualize! See the `visualize-data` vignette for instructions.
