---
title: "Importing and Using Local PurpleAir Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Importing and Using Local PurpleAir Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy = TRUE,
  eval = FALSE
)
```

```{r setup}
library(AirVizR)
library(magrittr)
```

## Inputs

Many functions rely on the same inputs, therefore default to the same defined objects in the environment.
It is highly recommended to specify inputs upfront in their own chunk to simplify functionality.
Various environment cleans are present in this document to showcase where and when these objects can be disgarded.

```{r inputs}
# Path to folder containing CSV files of interest
input_path <- "data-raw/local"

# Timezone in which the monitors are located
# Run OlsonNames() for a list of valid inputs
input_timezone <- "America/Los_Angeles"


#OPTIONAL: filter for specific locations, e.g. for STAR labs use "*STAR*" exactly. Use NULL otherwise
input_path_filter <- NULL


## Optional grouping settings (which will NOT impact raw data output, but can be useful when graphing!)

## DATE GROUPING
# Run date grouping? If set to "FALSE", the date inputs below will not matter (but will still appear in the environment)
run_date_grouping <- TRUE
# If TRUE, specify the groupings below as respective items in their lists. Dates that are not included in these ranges will be dropped.
# (i.e. the first item in each list is one "group", the second item in each list is another "group", etc.)
# Date grouping categories:
input_date_tags <- c("Before", "Independence Day", "After")
# Start dates (in a list, "YYYY-MM-DD")
input_date_starts <- c("2020-07-01", "2020-07-04", "2020-07-05")
# End dates (in a list, "YYYY-MM-DD")
input_date_ends <- c("2020-07-03", "2020-07-04", "2020-07-07")

## HOUR GROUPING
# Run hour grouping? If set to "FALSE", the hour inputs below will not matter (but will still appear in the environment)
run_hour_grouping <- TRUE
# If TRUE, specify the groupings below as respective items in their lists
# (i.e. the first item in each list is one "group", the second item in each list is another "group", etc.)
# Hour grouping categories:
input_hour_tags <- c("Morning", "Afternoon", "Evening", "Night")
# Start hours (in a list, using full hours in 24 hour format)
input_hour_starts <- c(5, 12, 17, 21)
# End hours will be automatically generated; no need to add "end" hours. This also means that no hours will be dropped.
# If you would like to drop specific hours, consider using the filter_df() function
```

---

## Loading Data

```{r reading_data}
local_meta <- read_local_meta(pattern = input_path_filter)

local_raw <- read_local_folder(pattern = input_path_filter)
```


## Wrangling Data


```{r clean_data}
local_full <- wrangle_local(local_raw, local_meta)

local_hourly <- apply_functions(local_full, by_day = TRUE, by_hour = TRUE)
local_daily <- apply_functions(local_full, by_day = TRUE, by_hour = FALSE)
local_diurnal <- apply_functions(local_full, by_day = FALSE, by_hour = TRUE)

if (run_date_grouping == TRUE){
  local_full <- local_full %>% apply_date_tags()
}
if (run_hour_grouping == TRUE) {
  local_full <- local_full %>% apply_hour_tags()
}
```

Pulling start and end dates (for other functions):

```{r start_end_dates}
if (exists("input_startdate") == FALSE) {
  input_startdate <- as.character(min((local_full %>% column_dt("date"))$date))
}
if (exists("input_enddate") == FALSE) {
  input_enddate <- as.character(max((local_full %>% column_dt("date"))$date))
}
```

```{r eclean_applying}
# OPTIONAL
# Cleaning the global environment
remove(run_date_grouping, run_hour_grouping)
remove(list = c(ls(pattern = "input_(hour|date)_")))
remove(list = c(ls(pattern = "input_(path|timezone)")))
# It is recommended to keep raw_meta and raw_data since it takes the longest to load,
# as well as the input start & end dates (since the diurnal set does not contain date information)
```

---

## Next steps

Now we're ready to visualize! See the [`visualize-data` vignette](visualize-data.html) for instructions.
